---
title: "Stat 435 Test 1"
author: "Brian Bruxvoort"
date: "2023-10-10"
output: pdf_document
---

```{r}
# Load the libraries and dataset
library(tidyverse)
library(dplyr)
shuttle <- read.csv("shuttle.csv")
```

# Question 1
## Part a
To estimate the population proportion of people who would say "yes", we use the sample population. The sample proportion is 344/1170 = 0.294 of people will say yes. We can assume the population proportion is this same value because as the sample proportion grows larger and larger it will eventually become the population proportion. So, the population proportion is 0.294. This is also our pi hat estimate.

## Part b
Ho : p = 0.5
Ha : p does not equal 0.5

z = ((344/1170)-0.5) / sqrt((0.5(1-0.5))/1170)
z = 6.22

Rejection Region = 1.96 when alpha = 0.05

Reject Ho, sufficient evidence to show that the proportion of people willing to accept cuts in their standard of living to protect the environment is different from 0.5. In other words, there is a statistically significant majority or minority in the population who would say "yes" to accepting cuts in their standard of living.

## Part c
```{r}
#Wald
w<-344 
n<-1170
alpha<-0.01 
pi.hat<-w/n 
var.wald<-pi.hat*(1-pi.hat)/n 
lower<-pi.hat-qnorm(p=1-alpha/2)*sqrt(var.wald) 
upper<-pi.hat+qnorm(p=1-alpha/2)*sqrt(var.wald) 
round(data.frame(lower,upper),4)

#Wilson
pi.tilde<-(w+qnorm(p=1-alpha/2)^2/2)/(n+qnorm(p=1-alpha/2)^2)

round(pi.tilde+qnorm(p=c(alpha/2,1-alpha/2))* sqrt(n)/(n+qnorm(p=1-alpha/2)^2)*sqrt(pi.hat* (1-pi.hat)+qnorm(1-alpha/2)^2/(4*n)),4)

#Agresti-Coull
var.ac<-pi.tilde*(1-pi.tilde)/(n+qnorm(p=1-alpha/2) ^2)  
round(pi.tilde+qnorm(p=c(alpha/2,1-alpha/2))* sqrt(var.ac),4)
```

The intervals are so similar and close because a 99% confidence interval for all three will mean they all need to be super accurate. When they all need to be super accurate, then the intervals will all be close because they are all going off the same pi hat.

# Question 2
## Part a
```{r}
c.table <- array(data = c (189, 104, 10845, 10933) , dim = c(2, 2),
dimnames = list("Group" = c("Surgery", "Radiation Therapy"),
"Cancer" = c("Controlled", "Not Controlled")))
c.table
```

```{r}
pi.hat.table <- c.table/rowSums(c.table)
pi.hat.table
```

```{r}
pi.hat1 <- pi.hat.table[1,1]
pi.hat2 <- pi.hat.table[2,1]

round(pi.hat1/pi.hat2, 4)

round(1/(pi.hat1/pi.hat2), 4) #inverted

alpha<-0.05
n1 <- sum(c.table[1,])
n2 <- sum(c.table[2,])
var.log.rr <- (1-pi.hat1) / (n1*pi.hat1) + (1-pi.hat2) / (n2*pi.hat2)
var.log.rr
```

```{r}
ci <- exp(log(pi.hat1/pi.hat2) + qnorm(p = c(alpha/2, 1-alpha/2)) * sqrt(var.log.rr))
round(ci, 4)

rev(round(1/ci, 4)) #inverted
```

The relative risk inverted is 0.5501, which means a patient is 45% more likely to not have the cancer controlled if they have radiation therapy done, as compared to having a surgery. Conversely a person that has the surgery done, is 45% more likely to have the cancer controlled as compared to the radiation therapy. With 95% confidence, we can say having the surgery will control the cancer between 30% and 57% more of the time than if the patient has radiation therapy. 

## Part b
```{r}
OR.hat <- c.table[1,1] * c.table[2,2] / (c.table[2,1] * c.table[1,2])
round(OR.hat, 4)

round(1/OR.hat, 4)
```

```{r}
alpha <- 0.05
var.log.or <- 1/c.table[1,1] + 1/c.table[1,2] + 1/c.table[2,1] + 1 /c.table[2,2]
OR.CI <- exp(log(OR.hat) + qnorm(p = c(alpha / 2, 1-alpha / 2)) * sqrt(var.log.or))
round(OR.CI, 4)

rev(round(1/OR.CI, 4))
```

The odds ratio says that the odds of the cancer being controlled when are 1.8 times as large than if the radiation therapy is done to try and control the cancer. The confidence interval states with 95% certainty that the odds of the cancer being controlled are between 1.44 and 2.33 times as large as compared to the radiation therapy.

## Part c
```{r}
chisq.test(x = c.table, correct = FALSE)
```

The Pearson Chi-Square test has a p-value of 0.1022 which is greater than the rejection region of p < 0.05. In this case the test says their is insufficient evidence to conclude the two treatments are different.

## Part d
If pi1 = pi2 then they equal 1, because pi1/pi2 = 1. If the odds ratio equals 1, then the odds of two events happening are equal. The equation to calculate the odds ratio is pi1(1-pi2)/pi2(1-pi1) if pi1 and pi2 are equal to each other, then the odds ratio must be 1. The relative risk is equal to pi1/pi2, which we already confirmed equal to 1, therefore the relative risk must be equal to 1.

# Question 3
## Part a
```{r}
mod.fit <- glm(formula = TD ~ temp, family = binomial(link = logit), data = shuttle)
mod.fit
```

```{r}
summary(mod.fit)
```

TD = 15.0429 - 0.2322temp

## Part b
```{r}
exp(mod.fit$coefficients[2])

1/exp(5*mod.fit$coefficients[2])
```

```{r}
beta.ci <- confint(object = mod.fit, parm = "temp", level = 0.95)
beta.ci
```

```{r}
as.numeric(rev(1/exp(beta.ci*5)))
```

With 95% confidence, the odds of a success change by an amount between 1.36 to 13.16 times for every 5 degree decrease in temperature.

Because the interval is entirely above 1, there is sufficient evidence that a 5 degree decrease in temperature increases the odds of a thermal distress.

## Part c
```{r}
linear.pred <- mod.fit$coefficients[1] + mod.fit$coefficients[2]*70
exp(linear.pred)/(1+exp(linear.pred))
as.numeric(exp(linear.pred)/(1+exp(linear.pred)))
```

## Part d
```{r}
curve(expr = exp(mod.fit$coefficients[1] + mod.fit$coefficients[2] * x) / (1 + exp(mod.fit$coefficients[1] + mod.fit$coefficients[2] * x)), col = "red" , xlim = c(50, 85), ylab = expression(hat(pi)), xlab = "Temp", main = "Estimated probability of thermal distress", panel.first = grid())
```

# Question 4

The answers for question 4 and its parts are in the attached document.
